---
title: "Stats Demo 1"
output: pdf_document
---

## Environment setup
```{r setup, echo=T, results='hide', error=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy.opts = list(width.cutoff = 60), tidy = TRUE)

library(Sleuth3)
library(reshape2)
library(ggplot2)
library(dplyr)
library(plotly)
#library(ggplot2)
```

# 1. Creating data
Creating data for test. Prices for CS course books in OSU bookstore.
Notice the difference between lists and Data frames.
```{r}
Prices <- c(99.34, 51.53, 20.45, 97.22, 61.89, 58.17, 61.63, 44.63, 96.69, 48.88)
Prices

typeof(Prices)

#check? data_frame(Prices)
is.data.frame(Prices)

#as trans the type
PricesDF <- as.data.frame(Prices)
colnames(PricesDF) <- ("Price")
PricesDF$Price
PricesDF

Genres <- factor(c("Science", "Engineering", "Art", "Science",
"Math", "Engineering", "Art", "Math", "Engineering", "Art"))
BookStore <- data.frame(Price = Prices, Genre = Genres)
BookStore


```

# 2. Working with data
Manipulating data frames is easier as each column becomes an object.
```{r}

```

# 3. Plots
`ggplot` provides a platform to plot various plots by changing the graph types. Refer to https://rstudio.com/wp-content/uploads/2015/03/ggplot2-cheatsheet.pdf for common commands.
The type of plot depends on the nature of the variable. Here we see how bar plots are better suited for discreet data:
```{r}

```

# 4. Distribution of Data
Sampling from any distribution, we can examine a histogram of the data:
```{r}

```

# 5. Central Limit Theorem
When repeated samples are drawn with a sample size~=30, the sample mean distribution approximately represents a normal distribution:
```{r}

```

# 6. One sample t-test
Test whether the mean price of your sample of CS books is significantly different from the mean price of all books in the OSU bookstore (population):
```{r}

```

# 7. Two sample t-test
What about when we compare against the mean price of all books on Amazon? We can compare using a similarly sized sample from Amazon. We then test if the mean prices are significantly different between the two samples:
```{r}

```

# 8. One and two-tail analysis

alternative: the alternative hypothesis. Allowed value is one of “two.sided” (default), “greater” or “less”.
var.equal: a logical variable indicating whether to treat the two variances as being equal. If TRUE then the pooled variance is used to estimate the variance otherwise the Welch test is used.
```{r}

```


# 9. Paired t-test
When there exists a natural pairing in the observations. For example, when comparing the prices of the same books in Amazon and OSU BookStore.

```{r}

```


# 10. Reporting

<!-- Conclusion statement goes here... -->



