---
title: "Stats Demo 1"
output: pdf_document
---

## Environment setup
```{r setup, echo=T, results='hide', error=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy.opts = list(width.cutoff = 60), tidy = TRUE)

library(Sleuth3) # example datasets from textbook, "The Statistical Sleuth - A Course in 
# Methods of Data Analysis (3rd Edition)"
library(reshape2) # for formatting and aggregation of data frames
library(ggplot2) # for creating graphs
library(dplyr) # for data manipulation and clean-up
library(plotly) # for creating interactive web graphics from ggplot2 graphs
library(knitr) # required for generating PDF output
library(modeest) # required for `mfv()` function
```

# 1. Creating data
Creating data for test. Prices for CS course books in OSU bookstore.
Notice the difference between lists (`Prices`) and dataframes (`PriceDf`).
```{r}
Prices <- c(99.34, 51.53, 20.45, 97.22, 61.89, 58.17, 61.63, 44.63, 96.69, 48.88)
Prices

# checking the type of the Prices variable
typeof(Prices)

# checking whether the Prices variable is a dataframe
is.data.frame(Prices)
```

```{r}
PriceDf <- as.data.frame(Prices)
colnames(PriceDf) <- ("Price")
PriceDf

# checking the type of the Prices variable
typeof(PriceDf)

# checking whether the Prices variable is a dataframe
is.data.frame(PriceDf)
```

Adding genres for each of the textbooks, and creating a new dataframe (`BookStore`) to hold both prices and genres.
```{r}
Genres <- factor(c('Science', 'Engineering', 'Art', 'Science', 'Math', 'Engineering', 'Art', 'Math', 'Engineering', 'Art'))

# the data.frame() function will infer column names from the names of variables provided to it, unless we explicitly set them ourselves
BookStore <- data.frame(Price=Prices, Genre=Genres)
BookStore
```

# 2. Working with Dataframes
Manipulating dataframes is easier (than using lists) as each column becomes an object.

Accessing columns within a data frame can be done using either `[`, `[[`, or `$` operators:
```{r}
# same result for read the data
BookStore["Price"]
BookStore[["Price"]]
BookStore$Price

```
Adding columns to an existing data frame can be done using either the `$` or `[` operators, or the `cbind` function:
```{r}
Published <- c(1921, 2004, 2016, 2001, 1958, 1996, 2010, 2002, 1999, 2014)

# add some data
BookStore$Published <- Published
# or [""]
# or 
BookStore

```

Bonus: To remove that duplicate `Published` column from our `BookStore` dataframe:
```{r}

BookStore <- BookStore[, !duplicated(names(BookStore))]
BookStore
```
Let's also look at a couple of descriptive statistics for the data in our dataframe:
```{r}
mean(BookStore$Price)
median(BookStore$Price)
mfv(BookStore$Price)
summary(BookStore$Price)

```

# 3. Replicating Data
Before we begin to analyze our data, we need to expand our data so that we have a reasonably large sample size. This is not appropriate in real-world data that is intended for publication, but is useful when setting up your analysis and visually inspecting the data.

We will resample our data out to 1000 textbooks that based on a normal distribution that is within 2 standard deviations of the mean:
```{r}
set.seed(1234)

ExpandePrices1 <- replicate(n=1, rnorm(1000, mean = mean(BookStore$Price), sd=2),simplify = FALSE[[1]])

# ExpandePrices <- rnorm(1000, mean = mean(BookStore$Price), sd=2)
# # trans to xx.xx
#ExpandePrices1 <- round(ExpandePrices1, 2)
#ExpandePrices1
# ExpandePrices
# 
# ExpandeGenres <- sample(unique(BookStore$Genre), replace = TRUE, size = 1000)
# ExpandeGenres
# 
# ExpandeBookStore <- data.frame(Prices = ExpandePrices, Genre = ExpandeGenres)
# ExpandeBookStore

```

What does the descriptive statistics of this larger dataset look like?
```{r}
mean(ExpandeBookStore$Prices)
median(ExpandeBookStore$Prices)
mfv(ExpandeBookStore$Prices)
summary(ExpandeBookStore$Prices)
```

# 4. Plots
`ggplot2` provides a platform to plot various plots by changing the graph types. Refer to https://rstudio.com/wp-content/uploads/2015/03/ggplot2-cheatsheet.pdf for common commands.
The type of plot depends on the nature of the variable.

## Visual inspections
First, lets look at a histogram of our smaller dataset (`BookStore`) and then a histogram of our larger dataset (`ExpandedBookStore`):
```{r}

hist(BookStore$Price)
hist(ExpandeBookStore$Prices)

```

## Customizing visualizations with `ggplot`
Neither of these plots uses the full power of `ggplot2`, so let's convert the larger dataset into a `ggplot` graph:
```{r}
ggplot(ExpandeBookStore, aes(x=Prices)) + geom_histogram(bins=30, color="black", file="grew") +labs(title = "OSU bookstore", x="Prices", y="Genre")


ggplot(data = ExpandeBookStore) + geom_point(
  mapping = aes(x = Prices, y = Genre))


```

## Different types of plots
Now we can begin to explore some other plot visualizations to get a feel for the distribution of our data. Perhaps we want to see what the distributions looks like when grouped by the genre:
```{r}
# Line plots

# ggplot(ExpandeBookStore, aes(x=Prices)) + geom_histogram(bins=30, color="black", file="grew") # +labs(title = "OSU bookstore", x="Princes", y="Genre")

ggplot(data = ExpandeBookStore) + geom_point(
  mapping = aes(x = Prices, y = Genre))

# Boxplots
ggplot(data = ExpandeBookStore) + geom_boxplot(
  mapping = aes(x = Prices, y = Genre, color = Genre)) + labs(title = "OSU Books", x="Prices", y="Genre")
```

For diversity, let's also look at the boxplots and scatter plots for a real dataset. The `iris` dataset is a built-in dataset in R that contains measurements on 4 different attributes (measured in centimeters) for 50 flowers from 3 different species:
```{r}

ggplot(iris,aes(x=Species, y=Sepal.Length)) + geom_boxplot()

ggplot(iris,aes(x=Species, y=Sepal.Length)) + geom_point()


```

# 5. Central Limit Theorem
When repeated samples are drawn with a sample size of approximately 30, the sample mean distribution approximately represents a normal distribution:
```{r}
# change the `n` parameter and re-run to see how the sample size changes the shape of the distribution
n <- 1000

sample <- rep(0, n)
for(i in 1:n){
	newSample <- rnorm(2)
	sample[i] <- mean(newSample)
}

hist(sample, 
	main=paste0('Histogram of Sample Medians (n = ', n, ')'), 
	xlab='Sample Median')
```